TacticToe: Learning to Prove with Tactics

### 2018-02-19 tactictoe_jv.tex
Sec 8

\paragraph{Instantiation of an Abstracted Tactic}
An abstracted tactic is not immediately applicable to a goal, since it contains
unspecified arguments. To apply an abstracted tactic $\hat{t}$, we first need
to
instantiate the placeholders inside $\hat{t}$. Because it is difficult to
build a general predictor that effective on each type of argument, we manually
design different argument predictors for each type. Those predictors are
given a goal $g$ as input and returns the best chosen argument for this
goal.

Our default algorithm relies on argument predictions for theorems with a 
radius of 16. It uses the produced list 
$\mathit{Predictions}^{\mathit{tactic}}_{16}(g)$ to replace all the 
placeholders 
in $\hat{t}$. 
List of theorems is a very common type in \holfour proofs and theorems contain 
enough information to be predicted accurately.
We also have positive results with term abstraction and plan to include it soon 
in our default algorithm.

\paragraph{Selection of Abstracted Tactics}
As abstracted tactics do not appear in human proofs, we need to find a way to
predict them so that they can contribute to the \tactictoe proof search.
A straightforward idea is to try $\hat{t}$ before $t$ during the proof search.
However, this risks doing unnecessary work as they may have a similar effect.
Therefore, we would like to decide
beforehand if one is better than the other.
In fact, we can re-use the orthogonalization module to do this task for us.
We add $\hat{t}$ to the competition, initially giving it the coverage of $t$.
If $\hat{t}$
wins, it is associated with $g$ and is included in the database of tactic
features and thus can be predicted during proof search.
After many orthogonalization competitions, the coverage of $\hat{t}$ may exceed 
the coverage of
$t$. At
this point, the coverage of $\hat{t}$ is estimated on its own and not inherited
from
$t$ anymore.

@@

The Table~\ref{theories} compares the success rates of re-proving for different
\holfour theories. \tactictoe outperforms \eprover on every
considered theory.
\eprover is more suited to deal with dense theories such as
\texttt{real} or \texttt{complex} where a lot of related theorems are available
and most proofs are usually completed by rewriting tactics. Thanks to its 
ability to re-use custom-built tactics, \tactictoe 
largely surpasses \eprover on theories relying on inductive terms and 
simplifications sets such as \texttt{arithmetic}, \texttt{list} 
and \texttt{f\_map}. Indeed, \tactictoe is able to recognize where and when to 
apply induction, a task at which ATPs are known to struggle with.

@@

We first investigate how \tactictoe and \eprover scales as the search time 
grows in Figure~\ref{fig:c1}.
In 10 seconds, \tactictoe solves 90 percent of the problem it can 
solve in 60 seconds. The analysis is a bit different for \eprover. Indeed,
we can clearly deduce from the bump at 30 seconds that \eprover is using at 
least two strategies and probably more. Strategies are a useful method to 
fight the exponential decline in the number of newly proven theorems over 
time. Therefore, integrating strategy scheduling in \tactictoe could be 
something to be experimented with. Overall, we recommend using \tactictoe and 
\eprover with a 10 seconds timeout. It is possible also to run them with a 
longer time limit, but we advise to do so in parallel to avoid waiting time.

@@

In order to appreciate the difficulty of the evaluation set from a human
perspective, the length distribution of human proofs in the
validation set is shown in Figure~\ref{fig:distrib}.  It is clear from the
graph that most of the human proofs are short. The proofs found by tactictoe 
follows about the same
distribution. This means that if \tactictoe finds a proof, there is about a 50
percent chance that it will be shorter than what a human would come up with.

@@

In Figure~\ref{fig:percentage}, we regroup the human proofs by their length to
see how well \tactictoe and \eprover performs on increased level of
difficulties. As expected, the longer the original proof is, the harder it is
for \tactictoe and \eprover to re-prove the theorem by their own mean.
The performance of \tactictoe is compelling with more than half of the theorem
that required a proof of length six being reproven. It is also consistently
better than \eprover for any proof length.

>>>>

\paragraph{Instantiation of an Abstracted Tactic}
An abstracted tactic is not immediately applicable to a goal, since it contains
unspecified arguments. To apply an abstracted tactic $\hat{t}$, we first need
to
instantiate the placeholders inside $\hat{t}$. Because it is difficult to
build a general predictor effective on each type of argument, we manually
design different argument predictors for each type. Those predictors are
given a goal $g$ as input and return the best predicted argument for this
goal.

Our default algorithm relies on argument predictions for theorems with a 
radius of 16. It uses the produced list 
$\mathit{Predictions}^{\mathit{tactic}}_{16}(g)$ to replace all the 
placeholders 
in $\hat{t}$. 
The type of theorem lists is a very common type in \holfour proofs and theorems contain 
enough information to be predicted accurately.
We also have positive results with term abstraction and plan to include it soon 
in our default algorithm.

\paragraph{Selection of Abstracted Tactics}
As abstracted tactics do not appear in human proofs, we need to find a way to
predict them so that they can contribute to the \tactictoe proof search.
A straightforward idea is to try $\hat{t}$ before $t$ during the proof search.
However, this risks doing unnecessary work as the two may perform similar steps.
Therefore, we would like to decide
beforehand if one is better than the other.
In fact, we can re-use the orthogonalization module to do this task for us.
We add $\hat{t}$ to the competition, initially giving it the coverage of $t$.
If $\hat{t}$
wins, it is associated with $g$ and is included in the database of tactic
features and thus can be predicted during proof search.
After many orthogonalization competitions, the coverage of $\hat{t}$ may exceed 
the coverage of
$t$. At
this point, the coverage of $\hat{t}$ is estimated on its own and not inherited
from
$t$ anymore.

@@

Table~\ref{theories} compares the re-proving success rates for different
\holfour theories. \tactictoe outperforms \eprover on every
considered theory.
\eprover is more suited to deal with dense theories such as
\texttt{real} or \texttt{complex} where a lot of related theorems are available
and most proofs are usually completed by rewriting tactics. Thanks to its 
ability to re-use custom-built tactics, \tactictoe 
largely surpasses \eprover on theories relying on inductive terms and 
simplification sets such as \texttt{arithmetic}, \texttt{list} 
and \texttt{f\_map}. Indeed, \tactictoe is able to recognize where and when to 
apply induction, a task at which ATPs are known to struggle with.

@@

We first investigate how \tactictoe and \eprover scale as the search time 
grows in Figure~\ref{fig:c1}.
In 10 seconds, \tactictoe solves 90 percent of the problems it can 
solve in 60 seconds. The analysis is a bit different for \eprover. Indeed,
we can clearly deduce from the bump at 30 seconds that \eprover is using at 
least two strategies and probably more. Strategies are a useful method to 
fight the exponential decline in the number of newly proven theorems over 
time. Therefore, integrating strategy scheduling in \tactictoe could be 
something to be experimented with. Overall, we recommend using \tactictoe and 
\eprover with a 10 seconds timeout. It is possible also to run them with a 
longer time limit, but we advise to do so in parallel to avoid waiting time.
\todo{its strange to ``advise'' in a paper?}

@@

In order to appreciate the difficulty of the evaluation set from a human
perspective, the length distribution of human proofs in the
validation set is shown in Figure~\ref{fig:distrib}.  It is clear from the
graph that most of the human proofs are short. The proofs found by \tactictoe 
follows a similar
distribution. This means that if \tactictoe finds a proof, there is about a 50
percent chance that it will be shorter than what a human would come up with.

@@

In Figure~\ref{fig:percentage}, we regroup the human proofs by their length to
see how well \tactictoe and \eprover performs on increased level of
difficulties. As expected, the longer the original proof is, the harder it is
for \tactictoe and \eprover to re-prove the theorem on their own.
The performance of \tactictoe is compelling with more than half of the theorems
that required a proof of length six being reproven. It is also consistently
better than \eprover for any proof length.
